---
title: "DataScienceMathWeek4HW"
author: "KarenWeigandt"
date: "July 23, 2015"
output: html_document
---

#Data Science Math Week 4 Homework

1. Using matrix operations, describe the solutions for the following family of equations:
x + 2y - 3z = 5
2x + y - 3z = 13
-x + y = -8

This can be written as Ax = B in matrix form 

See Word document for Matrix form
    
The determinant of A =
  1 x ((1 x 0) - (-3 x 1)) - 2 x ((2 x 0) - (-3)(-1)) + (-3)((2 x 1) -  1(-1))
  = 1 x (0 + 3) - 2 x (0 - 3) + (-3) x (2 + 1)
  = 1 x 3 - (2)(-3) + (-3)(3)
  = 3 + 6 - 9
  = 0
  
If the determinant of a matrix is zero, then the system of equations is linearly dependent, and there is no unique solution. 



2. Provide a solution for #1, using R functions of your choice.

```{r}
# Build matrix M

vec1 <- c(1, 2, -1, 2, 1, 1, -3, -3, 0)

M = matrix(vec1, nrow = 3)

# Calulate determinant

det(M)

```
3. Solve for AB by hand:

See Word document 



4. Solve AB from #3 using R functions of your choice.

```{r}
# Build matrix A

vec2 = c(4, -3, 0, -3, 5, 1)

A = matrix(vec2, nrow = 3)

A

# Build matrix B

vec3 = c(1, 3, 4, -2)

B = matrix(vec3, nrow = 2)

B

# Multiply

D <- A %*% B

print("Matrix D = Matrix A x Matrix B")

D

```


